Effects of below-ground space limitation on performance of Eucalyptus seedlings: Does photosynthesis really control growth?

### 2015-07-20 tables_latex/pve_manuscript.tex
added new folder with latek markdown for tables

\documentclass[a4paper]{article}

@@

%allows not italic greek letters
\usepackage{textgreek}

\begin{document}

@@

<<preload, echo=FALSE>>=
library(knitr)
opts_chunk$set(echo=FALSE)
@

@@

Alternatively, manipulation of plant tissue carbon sinks is often used to investigate the correlation of photosynthesis and growth. Tissue sink activity can restrict biomass production when limited by environmental or developmental constraints \citep{korner2003carbon}. This is because metabolic signaling networks, relaying information on \textit{C} and nitrogen (\textit{N}) status of different tissues, can down regulate photosynthetic activity \citep{paul2001sink}.  Whether this down regulation, via sink inhibition, is evident in woody species has been tested through fruit removal, girdling, and low temperatures. In these studies, down regulation of \textit{A} was frequently correlated to carbohydrate accumulation resulting from reduced tissue sink strength \citep{iglesias2002regulation,hoch2002altitudinal,urban2007girdling, haouari2013fruit}. However, reductions in \textit{A} were also attributed to biochemical limitations prior to carbohydrate accumulation \citep{nebauer2011photosynthesis}, irreversible photo-oxidative damage \citep{duan2008photosynthetic}, and stomatal limitation \citep{li2005photosynthesis}. These mixed results are not surprising as we still know little about the pathways in which plants achieve balance between assimilation, storage, and growth across temporal scales \citep{smith2007coordination}. As these manipulations likely impact source as well as sink activity simultaneously, affect water transport, are very extreme, or are specific to large annual fruiting sinks, they tell us little about source-sink coordination in ‘normal’ field-grown woody species.

An alternative experimental approach is to lower belowground \textit{C} sink strength in tree seedlings by manipulating rooting volume, by varying the container size. The advantage of this approach is it allows a large range of manipulations, can be easily compared to naturally planted seedlings and mimics natural conditions as seedlings compete for space or reach bedrock. Seedlings undergo many physiological and morphological changes in response to rooting volume, including biomass partitioning, \textit{A}, water relations, nutrient uptake and respiration \citep[and references therein]{nesmith1998effect}. Inadequate rooting volume may decrease carbon sink strength by progressively restricting root growth \citep{thomas1991root}.  Container size studies frequently have photosynthetic down-regulation, likely as a result of sink limitation \citep{arp1991effects, mcconnaughay1991physical, gunderson1994photosynthetic, sage1994acclimation, maina2002intra, ronchi2006growth}. A meta-analysis by \citet{poorter2012pot} concluded that \textit{A} is the process likely to be the strongest affected by pot size and may best explain the observed effect on biomass seen in the large number of studies where containers are used. This conclusion arises because plants grown in small containers are shown to accumulate leaf starch while having lower \textit{C} exchange and assimilate export rates \citep{robbins1988effect}. However, evidence in support for a trade-off between \textit{C} storage and growth in trees is, to date, inconclusive \citep{palacio2014does}. Based on these previous studies, using container size as a sink-strength manipulation can be used to empirically test how growth and \textit{A} are coordinated.

Our lack of knowledge regarding the allocation of \textit{C} in woody species is still a major obstable in understanding the cooridatnation between  \textit{A} and growth. Understanding the fate of assimilated \textit{C} is crucial when source or sink manipulations do not directly lead to equivalent changes in growth. As programmed plasticity in \textit{C} allocation naturally occur as plants growth and develop, this must be accounted for when evaluating induced variation in growth \citep{wright2002interpreting}. Thus, when \textit{C} allocation is altered by a treatment, it should be shown whether these allocation patterns differ from common sized untreated plants \citep{reich2002root, poorter2012biomass}. Only then can we test if functional balance preservation or optimal foraging occurs from changes in \textit{C} uptake. For example, seedlings in containers may shift allocation to leaves if root restriction occurs or alter fine root morphology to increase nutrient uptake. In woody species, shifts in allometric relationships of leaves, stems and roots can reveal if growth is ontogenetically constrained or actively adjusted when \textit{A} is affected. This then allows additional pools of \textit{C}, such as root exudation or changes in tissue respiration, to be accounted for and evaluated.

@@

%--------------------------------------------------------------------------------------------%
\section*{Tables}
% Load xtable for printing tables.
<<echo=FALSE, message=FALSE, warning=FALSE>>=
library(xtable)
pve_data <- read.csv("table/pve_table1.csv")

colnames(pve_data) <- c("Volume (L)", 
                        "Seedling~mass~(g)", 
                        "SLA\\textsubscript{TNC-free}~(m\\textsuperscript{2}~kg\\textsuperscript{-1})",
                        "Leaf~Starch~(\\%)", 
                        "Leaf~Sugars~(\\%)", 
                        "Leaf~Nitrogen~(\\%)",
                        "Root~Nitrogen~(\\%)",
                        "SRL~(cm~m\\textsuperscript{-1})",
                        "{Leaf~\\textdelta}\\textsuperscript{13}C~(\\text{\\textperthousand})")

phys_data <- read.csv("table/pve_table2.csv")
colnames(phys_data) <- c("Volume~(L)", 
                        "\\textit{A}\\textsubscript{max}", 
                        "\\textit{R}\\textsubscript{dark}",
                        "\\textit{J}\\textsubscript{max}",
                        "\\textit{Vc}\\textsubscript{max}",
                        "\\textit{g}\\textsubscript{s}",
                        "\\textit{g}\\textsubscript{1}")

@

%seedling data table
<<results='asis'>>=
datatab <- xtable(pve_data, label="table:Table1", caption="Responses of plant and leaf characterisitics of \\textit{Eucalyptus tereticornis} seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Seedling mass, SRL, root nitrogen and leaf \\textdelta\\textsuperscript{13}C values are from final harvest. Values of leaf starch, sugars, nitrogen and SLA represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.")

print(datatab, include.rownames=FALSE,sanitize.colnames.function=function(x){x},
      floating.environment='sidewaystable', caption.placement="top", 
      hline.after = getOption("xtable.hline.after", c(-1,0,7,nrow(datatab))))

@

@@

<<results='asis'>>=
phystab <- xtable(phys_data, label="table:Table2",caption="Responses of leaf level gas exchange parameters of \\textit{Eucalyptus tereticornis} seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Units for \\textit{A}\\textsubscript{max} and \\textit{R}\\textsubscript{dark} are {\\textmugreek}mol~m\\textsuperscript{-2}~s\\textsuperscript{-1} and \\textit{g}\\textsubscript{s} are mol~m\\textsuperscript{-1}~s\\textsuperscript{-1}, each at at 25$\\degree$C. Values of \\textit{A}\\textsubscript{max}, \\textit{g}\\textsubscript{s} and \\textit{g}\\textsubscript{1} represent overall means across measurement campaigns (n=6). \\textit{R}\\textsubscript{dark}, \\textit{J}\\textsubscript{max} and \\textit{Vc}\\textsubscript{max} values are means of two measurement campaigns at beginning and end of gas exchange measurements. Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.")

print(phystab, include.rownames=FALSE,sanitize.colnames.function=function(x){x},
      floating.environment='sidewaystable',caption.placement="top",
       hline.after = getOption("xtable.hline.after", c(-1,0,7,nrow(datatab))))
@

>>>>

\documentclass[a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

%allows not italic greek letters
\usepackage{textgreek}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

@@

Alternatively, tissue sink activity can restrict biomass production when limited by environmental or developmental constraints \citep{korner2003carbon}. This is because metabolic signaling networks, relaying information on \textit{C} and nitrogen (\textit{N}) status of different tissues, can down regulate photosynthetic activity \citep{paul2001sink}.  Whether this down regulation, via sink inhibition, exists in woody species has been tested through fruit removal, girdling, and low temperatures. In these studies, down regulation of \textit{A} was frequently correlated to carbohydrate accumulation resulting from reduced tissue sink strength \citep{iglesias2002regulation,hoch2002altitudinal,urban2007girdling, haouari2013fruit}. However, reductions in \textit{A} were also attributed to biochemical limitations prior to carbohydrate accumulation \citep{nebauer2011photosynthesis}, irreversible photo-oxidative damage \citep{duan2008photosynthetic}, and stomatal limitation \citep{li2005photosynthesis}. These mixed results are not surprising as we still know little about the pathways in which plants achieve balance between assimilation, storage, and growth across temporal scales \citep{smith2007coordination}. As these manipulations likely impact source as well as sink activity simultaneously, affect water transport, are very extreme, or are specific to large annual fruiting sinks, they tell us little about source-sink coordination in ‘normal’ field-grown woody species. This coordination is further confounded by the fact that under normal growing conditions, \textit{A} is not always correlated to photosynthate accumulation, as with fir needles \citep{little1973effect}. Unfortunately, there is still limited understating of the physiological roles of carbohydrates in photosynthetic regulation and the elements triggering the down-regulation process \citep{nebauer2011photosynthesis}.

An alternative approach is to lower belowground \textit{C} sink strength in tree seedlings by manipulating rooting volume, by varying the container size. The advantage of this approach is it allows a large range of manipulations, can be easily compared to naturally planted seedlings and mimics natural conditions as seedlings compete for space or reach bedrock. Seedlings undergo many physiological and morphological changes in response to rooting volume, including biomass partitioning, \textit{A}, water relations, nutrient uptake and respiration \citep[and references therein]{nesmith1998effect}. The rooting volume available to plants can decrease \textit{C} sink strength by reducing root growth. Container size studies frequently have photosynthetic down-regulation, likely as a result of sink limitation \citep{arp1991effects, mcconnaughay1991physical, gunderson1994photosynthetic, sage1994acclimation, maina2002intra, ronchi2006growth}. A meta-analysis by \citet{poorter2012pot} concluded that \textit{A} is the process likely to be the strongest affected by pot size and may best explain the observed effect on biomass seen in the large number of studies where containers are used. This conclusion arises because plants grown in small containers are shown to accumulate leaf starch while having lower \textit{C} exchange and assimilate export rates \citep{robbins1988effect}. However, evidence in support for a trade-off between \textit{C} storage and growth in trees is, to date, inconclusive \citep{palacio2014does}. Based on these previous studies, using container size as a sink-strength manipulation can be used to empirically test how growth and \textit{A} are coordinated.

Another barrier in understanding the coordination between \textit{A} and growth is a lack of knowledge regarding the allocation of \textit{C} in woody species.  Understanding the fate of assimilated \textit{C} is crucial when source or sink manipulations do not directly lead to equivalent changes in growth. As programmed plasticity in \textit{C} allocation naturally occur as plants growth and develop, this must be accounted for when evaluating induced variation in growth \citep{wright2002interpreting}. Thus, when \textit{C} allocation is altered by a treatment, it should be shown whether these allocation patterns differ from common sized untreated plants \citep{reich2002root, poorter2012biomass}. Only then can we test if functional balance preservation or optimal foraging occurs from changes in \textit{C} uptake. For example, seedlings in containers may shift allocation to leaves if root restriction occurs or alter fine root morphology to increase nutrient uptake. In woody species, shifts in allometric relationships of leaves, stems and roots can reveal if growth is ontogenetically constrained or actively adjusted when \textit{A} is affected. This then allows additional pools of \textit{C}, such as root exudation or changes in tissue respiration, to be accounted for and evaluated.

@@

%--------------------------------------------------------------------------------------------%
\section*{Tables}
% Load xtable for printing tables.

%seedling data table
% latex table generated in R 3.1.2 by xtable 1.7-4 package
% Sun Jul 12 11:12:17 2015
\begin{sidewaystable}[ht]
\centering
\caption{Responses of plant and leaf characterisitics of \textit{Eucalyptus tereticornis} seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Seedling mass, SRL, root nitrogen and leaf \textdelta\textsuperscript{13}C values are from final harvest. Values of leaf starch, sugars, nitrogen and SLA represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.} 
\label{table:Table1}
\begin{tabular}{lllllllll}
  \hline
Volume (L) & Seedling~mass~(g) & SLA\textsubscript{TNC-free}~(m\textsuperscript{2}~kg\textsuperscript{-1}) & Leaf~Starch~(\%) & Leaf~Sugars~(\%) & Leaf~Nitrogen~(\%) & Root~Nitrogen~(\%) & SRL~(cm~m\textsuperscript{-1}) & {Leaf~\textdelta}\textsuperscript{13}C~(\text{\textperthousand}) \\ 
  \hline
5 & 14.8 (1.82) a & 11.8 (0.32) a & 12.7 (0.97) b & 6.4 (0.28) a & 1.1 (0.02) a & 0.78 (0.04) ab & 39.1 (5.47) a & -30.1 (0.26) a \\ 
  10 & 20.0 (2.38) ab & 11.7 (0.31) a & 9.4 (0.75) ab & 6.7 (0.25) a & 1.3 (0.04) ab & 0.75 (0.02) a & 34.2 (5.83) a & -30.2 (0.25) a \\ 
  15 & 25.4 (2.49) ab & 12.7 (0.48) a & 7.3 (0.73) a & 7.2 (0.28) a & 1.4 (0.06) ab & 0.71 (0.02) a & 37.6 (4.63) a & -30.3 (0.36) a \\ 
  20 & 23.4 (1.63) ab & 11.8 (0.37) a & 9.5 (0.88) ab & 6.6 (0.26) a & 1.4 (0.05) ab & 0.76 (0.04) a & 45.3 (5.50) a & -29.7 (0.28) a \\ 
  25 & 30.4 (5.49) ab & 12.4 (0.40) a & 9.8 (0.71) ab & 6.9 (0.24) a & 1.3 (0.06) ab & 0.74 (0.02) a & 47.0 (7.10) a & -29.7 (0.25) a \\ 
  35 & 52.2 (9.55) b & 13.5 (0.46) ab & 9.8 (0.65) ab & 6.8 (0.22) a & 1.5 (0.08) b & 0.77 (0.03) ab & 50.6 (11.61) a & -30.6 (0.38) a \\ 
  Free & 174.5 (18.02) c & 15.1 (0.47) b & 6.8 (0.65) a & 7.4 (0.25) a & 2.4 (0.09) c & 0.9 (0.03) b & 43.7 (6.24) a & -30.0 (0.34) a \\ 
   \hline
Container Effect (P) & 0.001 & 0.001 & 0.039 & 0.128 & 0.001 & 0.015 & 0.662 & 0.458 \\ 
   \hline
\end{tabular}
\end{sidewaystable}

@@

% latex table generated in R 3.1.2 by xtable 1.7-4 package
% Sun Jul 12 11:12:17 2015
\begin{sidewaystable}[ht]
\centering
\caption{Responses of leaf level gas exchange parameters of \textit{Eucalyptus tereticornis} seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Units for \textit{A}\textsubscript{max} and \textit{R}\textsubscript{dark} are {\textmugreek}mol~m\textsuperscript{-2}~s\textsuperscript{-1} and \textit{g}\textsubscript{s} are mol~m\textsuperscript{-1}~s\textsuperscript{-1}, each at at 25$\degree$C. Values of \textit{A}\textsubscript{max}, \textit{g}\textsubscript{s} and \textit{g}\textsubscript{1} represent overall means across measurement campaigns (n=6). \textit{R}\textsubscript{dark}, \textit{J}\textsubscript{max} and \textit{Vc}\textsubscript{max} values are means of two measurement campaigns at beginning and end of gas exchange measurements. Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.} 
\label{table:Table2}
\begin{tabular}{lllllll}
  \hline
Volume~(L) & \textit{A}\textsubscript{max} & \textit{R}\textsubscript{dark} & \textit{J}\textsubscript{max} & \textit{Vc}\textsubscript{max} & \textit{g}\textsubscript{s} & \textit{g}\textsubscript{1} \\ 
  \hline
5 & 21.2 (0.9) a & 0.61 (0.04) a & 104.5 (3.3) a & 63.3 (2.5) a & 0.30 (0.01) a & 5.1 (0.1) bc \\ 
  10 & 22.3 (1.4) ab & 0.79 (0.06) a & 116.5 (7.5) a & 69.4 (4.7) a & 0.36 (0.01) ab & 5.4 (0.1) cd \\ 
  15 & 23.3 (1.2) ab & 0.70 (0.05) a & 125.4 (7.8) a & 80.8 (5.1) ab & 0.42 (0.01) ab & 5.8 (0.1) d \\ 
  20 & 26.1 (0.7) b & 0.73 (0.11) a & 131.5 (8.6) a & 82.1 (4.7) ab & 0.37 (0.01) ab & 4.9 (0.1) ac \\ 
  25 & 23.9 (0.9) ab & 0.53 (0.13) a & 132.8 (13.1) a & 79.0 (8.7) a & 0.30 (0.01) a & 4.5 (0.1) a \\ 
  35 & 25.0 (1.0) ab & 0.61 (0.04) a & 127.2 (6.1) a & 82.4 (3.6) a & 0.31 (0.01) a & 4.4 (0.2) a \\ 
  Free & 33.1 (0.7) c & 0.64 (0.07) a & 169.0 (8.2) b & 100.4 (3.3) b & 0.44 (0.01) b & 4.5 (0.1) ab \\ 
   \hline
Container Effect (P) & 0.001 & 0.039 & 0.001 & 0.002 & 0.001 & 0.079 \\ 
   \hline
\end{tabular}
\end{sidewaystable}
